---
title: "Analysis of HPV, SIFs and HIV networks"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

#LOAD REQUIRED PACKAGES

```{r Initialize Functions}
# Load Libraries
library(igraph)
library(SparseM)
library(Matrix)
library(plyr)
library(dynamicTreeCut)
library(tidyr)
library(clusterProfiler)
library(dplyr)
library(data.table)
library(scales)
library(stringr)
library(ggplot2)
library(ggsignif)
library(ggpubr)
library(ggraph)
library(ggrepel)
library(scatterpie)
library(VennDiagram)
library(janitor)
library(forcats)
library(stringi)
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
library(network)
library(plotrix)

# minimum cluster numbers desired
minclustersize = 7

```

#Load datasets Used in Many Chucks

```{r Load datasets}
#Viral protein-human interactions from AP-MS. Courtesy of Prof. Krogan Lab
Krogan_PPI = read.table("~/Codes_for_manuscript_outputs/InPut_datasets/Krogan_HIV_HPV_AP-MS_interactome.txt", header = T, sep = '\t')
 
# #Load secreted proteins from HPA
HPA_1 = read.table('~/Codes_for_manuscript_outputs/InPut_datasets/Protein_predicted_secreted_HPA.txt',header=T,sep='\t') #downloaded from HPA website
HPA_2 = read.table('~/Codes_for_manuscript_outputs/InPut_datasets/Protein_predicted_secretedmembrane_HPA.txt',header=T,sep='\t') #downloaded from HPA website
HPA_merged = rbind(HPA_1, HPA_2) # merged into one dataframe

#Load inflammatory factors secreted by patients on ART
HIV_inflamm_molec = fread("~/Codes_for_manuscript_outputs/InPut_datasets/Inflammatory_molecules_DEP_ARTvsHC.txt", header = T) #SIFs extracted from literature

```

#Source Packages Used in Many Chunks

```{r Source functions}
#For network propagation
source("~/Codes_for_manuscript_outputs/Source_functions/calc_laplacian.R")
source("~/Codes_for_manuscript_outputs/Source_functions/run_propagate_diffusion.R")
source("~/Codes_for_manuscript_outputs/Source_functions/run_propagate_perm_diffusion.R")
source("~/Codes_for_manuscript_outputs/Source_functions/calc_pvalues.R")

#For enrichment analysis
source('~/Codes_for_manuscript_outputs/Source_functions/do_enrichmap.R') #-- this is a customed method developed by Prof. Krogan Lab

```

#Perform Initial Gene Enrichment Analysis for HPV preys, SIFs, HPV preys and HIV-HPV overlapping preys

```{r Overall enrichment for Inflamm factors, HIV, HPV preys}
#Extract human proteins that interact with viral proteins
HIV_gene_list1 <- unique(Krogan_PPI$PreyGeneName[Krogan_PPI$pathogen=='HIV'])
HPV_gene_list1 <- unique(Krogan_PPI$PreyGeneName[Krogan_PPI$pathogen=='HPV'])
HIV_HPV_overlapping_interactors = intersect(HIV_gene_list1, HPV_gene_list1)

#select host proteins that interact ONLY with the HIV or HPV proteins
HIV_interactors = as.data.frame(c(setdiff(HIV_gene_list1, HIV_HPV_overlapping_interactors),setdiff(HIV_HPV_overlapping_interactors, HIV_gene_list1)))
colnames(HIV_interactors) = c("gene") #rename column

HPV_interactors = as.data.frame(c(setdiff(HPV_gene_list1, HIV_HPV_overlapping_interactors),setdiff(HIV_HPV_overlapping_interactors, HPV_gene_list1)))
colnames(HPV_interactors) = c("gene") #rename column

HIV_HPV_overlapping_interactors = as.data.frame(HIV_HPV_overlapping_interactors)
colnames(HIV_HPV_overlapping_interactors) = c("gene") #rename column

#Select SIFs
inflamm_prots_secreted = intersect(HIV_inflamm_molec$Genes,HPA_merged$Gene)
inflam_secreted = as.data.frame(inflamm_prots_secreted)
colnames(inflam_secreted) = c("gene")

#Prepare dataframe with preys and virus categories
HIV_interactors$Groups = rep("HIV PPIs", nrow(HIV_interactors))
HPV_interactors$Groups = rep("HPV PPIs", nrow(HPV_interactors))
HIV_HPV_overlapping_interactors$Groups = rep("HPV & HPV PPIs", nrow(HIV_HPV_overlapping_interactors))
inflam_secreted$Groups = rep("SIFs", nrow(inflam_secreted))

#Enrichment for the viral preys
preys_for_enrichment = rbind(HIV_interactors, HPV_interactors, HIV_HPV_overlapping_interactors, inflam_secreted)

#Perform enrichment
M=data.frame(genes=preys_for_enrichment$gene,group=preys_for_enrichment$Groups)
M=data.frame(genes=Top100sig_genes$nodes,group=Top100sig_genes$group)
cutHeight=0.8
topN=5

colnames(M)=c("genes","group")

pdf(file = "Figure_1F_All_inflamm_HIV_HPV_PPI_heatmap.pdf",width=8.38,height=5.58, useDingbats = FALSE)
enrich_table = do_enrichmap(M,cutHeight,topN,database="Pathways",flag_cluster_col=TRUE,flag_p=FALSE,path_to_databasefiles='~/Codes_for_manuscript_outputs/InPut_datasets/')
dev.off()

#save out enrichment table
fwrite(enrich_table, "Table_S1_inflamm_HIV_HPV_overlapping_preys.xls",sep='\t')
#----------------------------------------------------------------------------------------

#plot a venndiagram for all the preys
library(ggvenn)

venn <- list(`SIF` = c(inflamm_prots_secreted),
             `HPV` = c(HPV_gene_list1),
             `HIV` = c(HIV_gene_list1))
 
ggvenn(venn, c("SIF", "HPV", "HIV")) #use the values from this venndiagram for venn.plot below

#Use the values obtained from the preceding chunk for HIV, HPV and HIV-HPV overlapping preys for final venn diagrams
venn.plot <- draw.triple.venn(
area1 = 391,
area2 = 26,
area3 = 276,
n12 = 0,
n23 = 0,
n13 = 35,
n123 = 0,
category = c("HPV only PPIs", "SIFs", "HIV only PPIs"),
fill = c("#0073C2FF","forestgreen", "#CD534CFF"),
lty = "blank",
cex = 2,
cat.cex = 2,
cat.pos = c(285, 180, 105),
cat.dist = 0.09,
cat.just = list(c(0.5, 1), c(0.5, 1), c(0.5, 0)),
ext.pos = 30,
ext.dist = -0.05,
ext.length = 0.85,
ext.line.lwd = 2
);
grid.draw(venn.plot);
grid.newpage();

```

#PERFORM NETWORK PROPAGATION

```{r Prepare base network to use for network propagation}
# Network name
name = "network_PC_CORUM_Reactome_IntAct"
 
# # Load network from PC
KD_pathwaycommons = fread('D:/Charles_PhD_Stuff/R_scripts_notebooks/Special_files/PathwayCommons_Clean.txt') #Download from PathwayCommons

G_pathwaycommons = KD_pathwaycommons[grepl("CORUM",KD_pathwaycommons$Interaction_Data_Source)
                                     | grepl("Reactome",KD_pathwaycommons$Interaction_Data_Source)
                                     | grepl("IntAct",KD_pathwaycommons$Interaction_Data_Source),
                                    c("From","To")]
#save out base network
fwrite(G_pathwaycommons, "Table S9_Base_network_for_propagation.xls", sep = "\t") #save propagation outputs
#or directly load the G_pathwaycommons we prepared above
#G_pathwaycommons = fread('~/GitHub/Network_Modeling_HIV-HPV/InPut_dataset_and_functions/Base_network_from_PathwayCommons_clean.txt') #Table S7 in the manuscript

S = G_pathwaycommons

# Make final graph
S = unique(S)
S = S[as.logical(!rowSums(S=="")),]

# Make and refine graph (only need to do this once and only if needed for your network)
G = graph_from_data_frame(S, directed = FALSE, vertices = NULL)
G = decompose.graph(G, max.comps = 2)
G = G[[1]]
nodes_all=row.names(as.matrix(V(G)))
S_final = unique(as.data.frame(as_edgelist(G)))
colnames(S_final) = c("From","To")
length(unique(c(S_final$From,S_final$To)))

# Save out network
fwrite(S_final,paste0("~/Codes_for_manuscript_outputs/InPut_datasets/",name,".txt"),sep='\t')

```

#Load network
```{r Load network from previous chunk and make graph object}

# Network name
name = "network_PC_CORUM_Reactome_IntAct"

# Load in network
G_network = fread(paste0("~/Codes_for_manuscript_outputs/InPut_datasets/",name,".txt"),header=T,sep='\t')

# Select subset
#G_network = G_network[1:10000,] #comment after trial phase
G = graph_from_data_frame(G_network, directed = FALSE, vertices = NULL)
G = decompose.graph(G, max.comps = 2)
G = G[[1]]

# Continue
length(unique(c(G_network$From,G_network$To)))

# Make final graph
nodes_all = row.names(as.matrix(V(G)))

```

# Create Laplacian --- only run once; afterwards the enrire chunk should be commented

```{r Calculate Laplacian}
start_time <- Sys.time()
L_all = calc_laplacian(G)
#L_time_exp = L_all[[2]]
#L = L_all[[1]]
finish_time <- Sys.time()
print(finish_time-start_time)

save(L_all,file=paste0(name,"_invdenom.Rdata"))

```

#Only run once; afterwards the enrire chunk should be commented

```{r Do Network Propagation Multiply gene-wise propagation}
# Load exp Laplacian

load(paste0(name,"_invdenom.Rdata"))
L_time_exp = L_all[[2]]
L = L_all[[1]]

# # --------------------------------------------------------------------------------------------
#Load Gene List
#Extract human proteins that interact only with HIV or HPV
HIV_gene_list1 <- unique(Krogan_PPI$PreyGeneName[Krogan_PPI$pathogen=='HIV'])
HPV_gene_list1 <- unique(Krogan_PPI$PreyGeneName[Krogan_PPI$pathogen=='HPV'])
HIV_HPV_overlapping_interactors = intersect(HIV_gene_list1, HPV_gene_list1)

#First gene list
HIV_gene_list = c(setdiff(HIV_gene_list1, HIV_HPV_overlapping_interactors),setdiff(HIV_HPV_overlapping_interactors, HIV_gene_list1))

#Second gene list
HPV_gene_list = c(setdiff(HPV_gene_list1, HIV_HPV_overlapping_interactors),setdiff(HIV_HPV_overlapping_interactors, HPV_gene_list1))

#Third gene list
#Use secreted inflammatory molecules from literature
inflamm_prots_secreted = intersect(HIV_inflamm_molec$Genes,HPA_merged$Gene)

# # Make into a dataframe
D = data.frame(nodes = nodes_all)
D$HIV_preys = as.logical(match(D$nodes,HIV_gene_list,nomatch=FALSE))
D$HPV_preys = as.logical(match(D$nodes,HPV_gene_list,nomatch=FALSE))
D$SIFs = as.logical(match(D$nodes, inflamm_prots_secreted,nomatch = FALSE))

# # --------------------------------------------------------------------------------------------
#Run Propagation

for (i in 1:(ncol(D)-1)){
#for (i in 1:1){

  nodes2label = D$nodes[D[,i+1]]
  nodes2labelvalues = matrix(1,ncol=1,nrow=length(nodes2label)) #Label with 1, everything else will be labeled 0.

  # Run baseline propagation
  xss_t = run_propagate_diffusion(G,L_time_exp,nodes2label,nodes2labelvalues)
  xss_t_1 = xss_t[[1]]
  x0_1 = xss_t[[2]]

  # Run Propagate Permutation
  xss_perm_mat_1 = run_propagate_perm_diffusion(G,L_time_exp,nodes2label,nodes2labelvalues,num_perms=10000,flag_PermuteInLabeled=FALSE)

  # Calculate p-values
  Y_1 = calc_pvalues(G, xss_t=xss_t_1, xss_perm_mat=xss_perm_mat_1, x0=x0_1)

  # Create data frame
  Y_1$dataset = colnames(D)[i+1]

  if (i ==1){
    Y_final = Y_1
    Y_perm_final = list()
    Y_perm_final[[colnames(D)[i+1]]] = xss_perm_mat_1
  } else {
    Y_final = rbind(Y_final,Y_1)
    Y_perm_final[[colnames(D)[i+1]]] = xss_perm_mat_1
  }
}
fwrite(Y_final,"output_propagations_diffusion_individual.txt",sep="\t")
#save(Y_perm_final, file = "Y_perm_final.Rdata")

# # --------------------------------------------------------------------------------------------
# Data Integration
xss_t_combined = aggregate(Y_final$xss_t,
                           by=list(nodes=Y_final$nodes),
                           FUN = prod)
xss_t_combined = xss_t_combined[match(nodes_all,xss_t_combined$nodes),]

xss_perm_mat_combined = Reduce("*", Y_perm_final)

Y_combined = calc_pvalues(G, xss_t=xss_t_combined$x, xss_perm_mat=xss_perm_mat_combined)
fwrite(Y_combined,"output_propagations_combined.txt",sep="\t")

```

#Induce a subnetwork from either HIV-HPV or HPV-SIF propagation outputs

```{r Extract significant genes and induce a graph}
# Select significant nodes for two propagations
Y_final = fread("~/Codes_for_manuscript_outputs/InPut_datasets/output_propagations_diffusion_individual.txt")
Y.wide = reshape2::dcast(Y_final,nodes~dataset,value.var="pvals")
fwrite(Y_final, "Table 6_Network_propagation_scores.xls", sep = "\t") #save propagation outputs

#SIFs and HPV preys - run this to for SIFs-HPV propagation outputs ---Uncomment lines 229:337 to run HPV and HIV analysis
Y.wide$sig[rowSums(Y.wide[,c("HPV_preys","Secreted_inflamm")]) < 0.05]="YES" #convergence of HPV preys vs SIFs
Y.wide$sigsum = rowSums(Y.wide[c("HPV_preys","Secreted_inflamm")]) #calculate rowsum
Y.wide_clean = subset(Y.wide, sig=="YES")
sig_genes = Y.wide_clean$nodes
sig_genes = intersect(sig_genes,as_ids(V(G)))

#HPV and HIV preys ---Uncomment lines 218:223 to run HPV and HIV analysis
# Y.wide$sig[rowSums(Y.wide[,c("HIV_preys","HPV_preys")]) < 0.05]="YES" #convergence of HPV vs HIV
# Y.wide$sigsum = rowSums(Y.wide[c("HIV_preys","HPV_preys")]) #calculate rowsum
# Y.wide_clean = subset(Y.wide, sig=="YES")
# 
# prop_df_sorted = arrange(Y.wide_clean, sigsum)
# selected_sig_genes = as.data.frame(prop_df_sorted[1:500, "nodes"]) #select top 500 most significant genes
# colnames(selected_sig_genes) = c("nodes")
# sig_genes = selected_sig_genes$nodes
# sig_genes = intersect(sig_genes,as_ids(V(G)))

#Additional analysis: Enrichment of sig genes from HIV, HPV and SIFs propagation
#HIV propagation
Y.wide_HIV_sig = subset(Y.wide, HIV_preys<=0.05)
HIV_sig_sort = arrange(Y.wide_HIV_sig, HIV_preys)
Top500_sig_HIV_genes = as.data.frame(HIV_sig_sort[1:500, "nodes"]) #select top 500 most significant genes
colnames(Top500_sig_HIV_genes) = c("nodes")
Top500_sig_HIV_genes$group = rep("HIV", nrow(Top500_sig_HIV_genes))

#HPV propagation
Y.wide_HPV_sig = subset(Y.wide, HPV_preys<=0.05)
HPV_sig_sort = arrange(Y.wide_HPV_sig, HPV_preys)
Top500_sig_HPV_genes = as.data.frame(HPV_sig_sort[1:500, "nodes"]) #select top 500 most significant genes
colnames(Top500_sig_HPV_genes) = c("nodes")
Top500_sig_HPV_genes$group = rep("HPV", nrow(Top500_sig_HPV_genes))
 
# #SIFs propagation
Y.wide_SIF_sig = subset(Y.wide, Secreted_inflamm<=0.05)
SIF_sig_sort = arrange(Y.wide_SIF_sig, Secreted_inflamm)
Top500_sig_SIF_genes = as.data.frame(SIF_sig_sort[1:500, "nodes"]) #select top 500 most significant genes
colnames(Top500_sig_SIF_genes) = c("nodes")
Top500_sig_SIF_genes$group = rep("Hi-SIFs", nrow(Top500_sig_SIF_genes))

#Merge all the genes
Top100sig_genes = rbind(Top500_sig_HIV_genes,Top500_sig_HPV_genes,Top500_sig_SIF_genes)

#sig_genes = Top500_sig_HIV_genes$nodes
sig_genes = Top500_sig_HPV_genes$nodes
#sig_genes = Top500_sig_SIF_genes$nodes
sig_genes = intersect(sig_genes,as_ids(V(G)))

# Make subgraph
SN = induced_subgraph(G, sig_genes)

# Select only fully connected components of subnetwork
SN = igraph::simplify(SN)
components = igraph::clusters(SN, mode="weak")
biggest_cluster_id = which.max(components$csize)
vert_ids = V(SN)[components$membership == biggest_cluster_id]
SN = igraph::induced_subgraph(SN, vert_ids)

```

# Cluster Resulting Significant Propagated Network into Subnetwork Clusters

```{r Clustering based on Network Only}
# Perform initial clustering
cc = cluster_walktrap(SN, steps = 10) %>% as.hclust()
nodes_1 = cc$labels

# plot(cc)
cc_dist=cophenetic(cc) # calculates the distance until two nodes form into one branch.
cc_dist = as.matrix(cophenetic(cc))/max(as.matrix(cophenetic(cc))) # normalize based on max height
cc_dist_mat = as.matrix(cc_dist)
ct = cutreeDynamic(cc, cutHeight = NULL, minClusterSize = minclustersize, deepSplit = 4, method = 'hybrid', distM = cc_dist_mat)
cc_in = hclust(as.dist(cc_dist_mat))
ct_counts = plyr::count(ct)
node_table = data.frame(nodes=cc_in$labels,clusters=ct)

# Create new subnetwork object of cluster pieces
n=0
for (i in unique(ct)){
  SN_temp = induced_subgraph(SN,which(ct==i))
  #print(length(V(SN_temp)))
  if (n == 0){
    SN_final = SN_temp
    n=1
  } else {
    SN_final = igraph::union(SN_final,SN_temp)
  }

}
edges = as_edgelist(SN_final, names = TRUE)

#Plot
E(SN_final)$weight=1E-15
co = layout_with_fr(SN_final)
plot(SN_final, vertex.color = ct, vertex.size=2, vertex.border=F, layout = co)
pdf(file="network_clusters_layout.pdf",width=20,height=20)
plot(SN_final,vertex.size=1,vertex.label=NA, vertex.frame.width = 1E-8, vertex.color = node_table$clusters[match(as_ids(V(SN_final)),node_table$nodes)])
dev.off()

#Save out network edges file with cluster annotation
#fwrite(as.data.frame(edges),file="HPV_Inflamm_edges.txt",sep='\t') # for HPV-Inflamm 
#fwrite(as.data.frame(edges),file="HPV_HIV_edges.txt",sep='\t') # for HPV-HIV
#fwrite(as.data.frame(edges),file="HIV_only_edges.txt",sep='\t') # for HIV only
fwrite(as.data.frame(edges),file="HPV_only_edges.txt",sep='\t') # for HPV only
#fwrite(as.data.frame(edges),file="Hi-SIFs_only_edges.txt",sep='\t') # for Hi-SIFs only

#Save out nodes file
x0_table = dcast(Y_final,nodes~dataset,value.var="x0")
node_table = cbind(node_table,x0_table[match(node_table$nodes,x0_table$nodes),2:ncol(x0_table)])
#fwrite(node_table,file="HPV_HIV_nodes.txt",sep='\t') # for SIF-HPV 
#fwrite(node_table,file="HPV_inflamm_nodes.txt",sep='\t') # for HPV-HIV
#fwrite(node_table,file="HIV_only_nodes.txt",sep='\t') # for HIV only
fwrite(node_table,file="HPV_only_nodes.txt",sep='\t') # for HPV only
#fwrite(node_table,file="Hi-SIFs_only_nodes.txt",sep='\t') # for Hi-SIFs only

```

# Perform Gene Set Enrichment Analysis Per Subnetwork Cluster

```{r Gene set enrichment anaysis}
# Run enrichment
#M=fread('~/Codes_for_manuscript_outputs/HPV_inflamm_nodes.txt')[,c("nodes","clusters")] #Uncomment if runnung HPV-Inflamm propagated outputs
#M=fread('~/Codes_for_manuscript_outputs/HPV_HIV_nodes.txt')[,c("nodes","clusters")] #Uncomment if runnung HPV-HIV propagated outputs
#M=fread('~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/HIV_only_nodes.txt')[,c("nodes","clusters")] #Uncomment if runnung HIV only propagated outputs
M=fread('~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/HPV_only_nodes.txt')[,c("nodes","clusters")] #Uncomment if runnung HPV only propagated outputs
#M=fread('~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/Hi-SIFs_only_nodes.txt')[,c("nodes","clusters")] #Uncomment if runnung Hi-SIFs only propagated outputs

M = M[!grepl("_",M$nodes),]
colnames(M)=c("genes","group")

# Pathways
#pdf(file = "Figure 2C-Inflamm_HPV_overall_enrichment_pathway_heatmap.pdf",width=8.2,height=4.85,useDingbats = FALSE) # for HPV-inflamm
#pdf(file = "Figure S2B-HIV_HPV_enrichment_pathway_heatmap.pdf",width=14.11,height=5.61,useDingbats = FALSE) # for HPV-HIV
pdf(file = "Figure 2G1-HIV_only_enrichment_pathway_heatmap.pdf",width=7.82,height=3.85,useDingbats = FALSE) # for HIV only propagation
enrich_table_2 = do_enrichmap(M,cutHeight=0.8,topN=1,database="Pathways",flag_cluster_col=TRUE,flag_p=FALSE,
                            path_to_databasefiles= '~/Codes_for_manuscript_outputs/InPut_datasets/') #download from Msigdb
dev.off()

#save out
#fwrite(enrich_table_2,"Table-S2-HPV-SIFs_enrichment_table_pathways.xls",sep="\t")
#fwrite(enrich_table_2,"Table-S3-HIV_HPV_enrichment_table_pathways.xls",sep="\t")
#fwrite(enrich_table_2,"Table-S2-HIV_only_enrichment_table_pathways.xls",sep="\t")
#fwrite(enrich_table_2,"Table-S2-HPV_only_enrichment_table_pathways.xls",sep="\t")
fwrite(enrich_table_2,"Table-S2-Hi-SIFs_only_enrichment_table_pathways.xls",sep="\t")

#Enrichment plot for HIV only propagation
#load data
propag_enrichment_output_HIV = fread("~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/Table-S2-HIV_only_enrichment_table_pathways.xls")
enrichment_output_HIV = propag_enrichment_output_HIV
D_HIV = aggregate(enrichment_output_HIV$p.adjust,
by = list(enrichment_output_HIV$group),
FUN = function(x){min(x,na.rm=T)})
colnames(D_HIV) = c("group", "p.adjust")
D_HIV_fin = D_HIV %>%
  left_join(enrichment_output_HIV, by="p.adjust")
D_HIV_fin = subset(D_HIV_fin, !duplicated(D_HIV_fin[,1])) # delete any duplicates
D_HIV_fin = subset(D_HIV_fin, !duplicated(D_HIV_fin[,5])) # delete any duplicates

#subset columns of interest
D_HIV_fin_sub = D_HIV_fin %>% dplyr::select(Description, p.adjust)
D_HIV_fin_sub$group = rep("HIV", nrow(D_HIV_fin_sub))

propag_enrichment_output_HPV = fread("~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/Table-S2-HPV_only_enrichment_table_pathways.xls")
enrichment_output_HPV = propag_enrichment_output_HPV
D_HPV = aggregate(enrichment_output_HPV$p.adjust,
by = list(enrichment_output_HPV$group),
FUN = function(x){min(x,na.rm=T)})
colnames(D_HPV) = c("group", "p.adjust")
D_HPV_fin = D_HPV %>%
  left_join(enrichment_output_HPV, by="p.adjust")
D_HPV_fin = subset(D_HPV_fin, !duplicated(D_HPV_fin[,1])) # delete any duplicates
D_HPV_fin = subset(D_HPV_fin, !duplicated(D_HPV_fin[,5]))

#subset columns of interest
D_HPV_fin_sub = D_HPV_fin %>% dplyr::select(Description, p.adjust)
D_HPV_fin_sub$group = rep("HPV", nrow(D_HPV_fin_sub))

propag_enrichment_output_SIF = fread("~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/Table-S2-Hi-SIFs_only_enrichment_table_pathways.xls")
enrichment_output_SIF = propag_enrichment_output_SIF
D_SIF = aggregate(enrichment_output_SIF$p.adjust,
by = list(enrichment_output_SIF$group),
FUN = function(x){min(x,na.rm=T)})
colnames(D_SIF) = c("group", "p.adjust")
D_SIF_fin = D_SIF %>%
  left_join(enrichment_output_SIF, by="p.adjust")
D_SIF_fin = subset(D_SIF_fin, !duplicated(D_SIF_fin[,1])) # delete any duplicates
D_SIF_fin = subset(D_SIF_fin, !duplicated(D_SIF_fin[,5])) # delete any duplicates

#subset columns of interest
D_SIF_fin_sub = D_SIF_fin %>% dplyr::select(Description, p.adjust)
D_SIF_fin_sub$group = rep("Hi-SIFs", nrow(D_SIF_fin_sub))

#Prepare dataframe for heatmap
D_three_merged = rbind(D_HIV_fin_sub,D_HPV_fin_sub,D_SIF_fin_sub)

#Clean the dataset to remove outliers
D_three_merged$inv_log = -log10(D_three_merged$p.adjust)
D_three_merged_mut = subset(D_three_merged, inv_log < 200)

# #wide format
D.wide = reshape2::dcast(D_three_merged_mut,Description~group,value.var="inv_log")
D.wide_clean = D.wide %>% replace(is.na(.), 0)
# 
row.names(D.wide_clean) <- D.wide_clean$Description
D.wide_clean$Description <- NULL

#Convert dataframes into matrix
hmp <- data.matrix(D.wide_clean)

col_2 = colorRamp2(seq(min(hmp), max(hmp), length = 3), c("gray100", "hotpink", "hotpink4"))

#overall mut data
hmp_1 = Heatmap(hmp, name = "-log10(p.adjust)", col = col_2, clustering_method_rows = "ward.D2", column_names_gp = gpar(fontsize = 10), row_names_gp = gpar(fontsize=10), rect_gp = gpar(col = "gray48", lwd = 1),
              heatmap_legend_param = list (
                direction = "horizontal",
                title_position = "topcenter"))

pdf(file = "Figure 2G_enrichment plot.pdf", width = 4.5, height = 8.94, useDingbats = FALSE)
print(hmp_1)
dev.off()

#Plot and print output the enrichment plot
ppath = ggplot(D_three_merged_mut, aes(x = reorder(x=Description, -log10(p.adjust)), y = -log10(p.adjust), fill = group,))+
geom_bar(stat="identity", width = 0.7) +
  #facet_grid(.~group, scales="free_x")+
#scale_fill_gradient2(low = "blue3", mid = "darkviolet", high = "red1", midpoint = 12) +
  theme_bw()+
coord_flip()+
  xlab("")+
  #geom_text(aes(label = invlog),position = position_dodge(width = 0.7), vjust = 1, color = "black", size = 2.5)+
  theme_bw()+
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
      )+
    theme(strip.background = element_rect(fill="white"))

pdf(file = "Figure 2G_enrichment plot_HIVonly.pdf", width = 8.25, height = 3.82, useDingbats = FALSE)
print(ppath)
dev.off()

#Select top 20 terms for visualization
# propag_enrichment_output_df_sorted = arrange(propag_enrichment_output,p.adjust)
# propag_enrichment_output_1 = subset(propag_enrichment_output_df_sorted, !duplicated(propag_enrichment_output_df_sorted[,3])) # delete any duplicates
# propag_enrichment_output1 = propag_enrichment_output_1[1:35,] # select the top 20 terms
# propag_enrichment_output1$Description = as.data.frame(tolower(propag_enrichment_output1$Description))
# 
# #Plot and print output the enrichment plot
# p = ggplot(propag_enrichment_output1, aes(x = reorder(x=Description, -p.adjust), y = -log10(p.adjust), fill = Count))+
# geom_bar(stat="identity", width = 0.7) +
# scale_fill_gradient2(low = "blue3", mid = "darkviolet", high = "red1", midpoint = 20) +
#   theme_bw()+
# coord_flip()
# 
# #seve out as pdf
# pdf(file = "Figure 2D_PC_withselfheat_propagation_Enrichment_plot.pdf", width = 10.97, height = 5, useDingbats = FALSE)
# print(p)
# dev.off()

```

#Visualize Propagated HPV PPIs and SIFs Network Subnetwork Clusters

```{r Add information to the nodes and save out clusters ---Uncomment if running SIF/HPV}
# # Initialize nodes and edges tables
# edges = fread("~/Codes_for_manuscript_outputs/HPV_Inflamm_edges.txt") # for Inflamm-HPV
# edges = data.table(from = edges[,1], to = edges[,2], type = "PC")
# nodes_in_edges_all = unique(c(edges$from,edges$to))
# nodes_df_clusters = data.frame(nodes = nodes_in_edges_all, cluster = ct[match(nodes_in_edges_all,cc_in$labels,nomatch=F)])
# 
# #select secreted inflammatory factors
# inflamm_prots_secreted = inflam_secreted$gene
# 
# #Prepare network edgelist
# n=0
# G_edges = as.data.table(as_edgelist(G))
# for (i in unique(nodes_df_clusters$cluster)){
#   nodes_cluster = as.character(nodes_df_clusters$nodes[nodes_df_clusters$cluster==i])
# 
#   # Prepare subnetwork
#   G_cluster = G_edges[as.logical(match(G_edges$V1,nodes_cluster,nomatch=F)) & as.logical(match(G_edges$V2,nodes_cluster,nomatch=F)),]
#   G_cluster$type = "PC"
#   G_cluster$cluster = i
# 
#     if (n==0){
#     G_Final_Subnetworks = G_cluster
#     n=1
#     } else {
#     G_Final_Subnetworks = rbind(G_Final_Subnetworks,G_cluster)
#     }
# 
#   }
# 
# #save out edge list
# fwrite(G_Final_Subnetworks,"Y_final_HIV_HPV_PPI_network_edges.txt",sep='\t') # network edgelist
# 
# #make node files
# nodes = data.frame(nodes = unique(c(as.character(G_Final_Subnetworks$V1),as.character(G_Final_Subnetworks$V2))))
# nodes$nodes_raw = gsub("_.*","",nodes$nodes)
# 
# # Add clusters to final nodes file
# nodes$cluster = nodes_df_clusters$cluster[match(nodes$nodes_raw,nodes_df_clusters$nodes)]
# nodes$cluster_all = nodes$cluster
# nodes$cluster_all[grepl("_",nodes$nodes)] = as.numeric(gsub(".*_","",nodes$nodes[grepl("_",nodes$nodes)]))
# 
# #Flag SIFs
# flag_inflamm_factors = as.logical(match(nodes$nodes_raw,inflamm_prots_secreted,nomatch=F))
# nodes$flag_inflamm_factors[flag_inflamm_factors]="Inflamm_factors"
# 
# #add -log pvalues from network propagation
# nodes$propagation_neglog10pvalue = -log10(Y_final$pvals[match(nodes$nodes,Y_final$nodes)])
# nodes$propagation_neglog10pvalue[!is.finite(nodes$propagation_neglog10pvalue)]=0
# 
# #Save out the node table
# fwrite(nodes,"Y_final_HIV_HPV_PPI_network_nodes_with_flags.txt",sep='\t') # network nodes
# 
# scale = 1.5
# # Load data
# G_Final_Subnetworks = fread("~/Codes_for_manuscript_outputs/Y_final_HIV_HPV_PPI_network_edges.txt")
# 
# # Load nodes
# nodes = fread("~/Codes_for_manuscript_outputs/Y_final_HIV_HPV_PPI_network_nodes_with_flags.txt")
# 
# # All together
# G_edges = G_Final_Subnetworks
# g = graph_from_data_frame(G_edges)
# layout = create_layout(g, layout = 'fr',circular=FALSE)
# layout$node_clusters = nodes$cluster_all[match(nodes$nodes,layout$name,nomatch=F)]
# node_types = rep("Human protein",nrow(layout))
# layout$node_types = node_types
# Y_final[, 2][Y_final[, 2] == 0] <- 0.00001 #replace the pvals = 0 with smaller values
# layout$propagation_sig = -log10(Y_final$pvals[match(layout$name,Y_final$nodes)])
# layout$propagation_sig[is.na(layout$propagation_sig)]=2
# layout[,9] <- rescale(layout[,9],to=c(0.8,1.3))
# 
# # Make text annotation table
# cmp = components(g)
# layout$node_components = cmp$membership
# cmps = unique(layout$node_components)
# xs = rep(NA,length(cmps))
# ys = rep(NA,length(cmps))
# cluster_name = rep(NA,length(cmps))
# for (i in 1:length(cmps)){
#   inds = as.logical(match(layout$node_components,cmps[i],nomatch=F))
#   xs[i] = mean(layout$x[inds])
#   ys[i] = max(layout$y[inds])
#   cluster_name[i]=unique(layout$node_clusters[inds])
# }
# 
# #ys = ys
# node_cmp_table = data.frame(components = cmps, xs = xs, ys = ys,
#                             cluster_name = cluster_name, cluster_labels = paste("C",cluster_name,sep=""))
# 
# # Define node size
# node_sizes_pies = layout$node_types
# node_sizes_pies[node_sizes_pies=="Human protein"] = 1.5/13
# layout$node_sizes_pies = as.numeric(node_sizes_pies)
# 
# # Create pie mat
# data_mat = nodes[,c("flag_inflamm_factors")]
# data_mat[data_mat=="Inflamm_factors"]=TRUE
# data_mat[data_mat==TRUE]=0.25
# data_mat$flag_inflamm_factors = as.numeric(data_mat$flag_inflamm_factors)
# data_mat[is.na(data_mat)]=0
# data_mat_in = cbind(layout,data_mat)
# data_mat_in$node_sizes_pies = layout$propagation_sig * layout$node_sizes_pies
# data_mat_in$flag_other = rep(0,nrow(data_mat_in))
# data_mat_in$flag_other[(data_mat_in[,c("flag_inflamm_factors")])==0]=0.25
# color_mat = c("red", "darkgrey")
# 
# #plot the network
# p = ggraph(layout) +
#     geom_edge_link(aes(color=type, width=1)) +
#     scale_edge_width(range = c(0.3, 0.3))+
#     scale_edge_color_manual(values = c("PC"="grey")) +
#     scale_color_manual(values = c("blue", "red", "green","darkgrey")) +
#     scale_shape_manual(values = c(20)) +
#     scale_size_manual(aes(values = propagation_sig*0.5)) +
#     geom_text_repel(aes(x=x,y=y,label=name), point.padding = 0.1,box.padding = 0.15, force_pull=0.1,
#                     max.overlaps = Inf,size=1.5,segment.linetype = 'dotted',segment.size=0.2,color="black") +
#     theme_classic() +
#     theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),
#           axis.ticks=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank()) +
#     geom_scatterpie(aes(x=x, y=y, r=node_sizes_pies),data = data_mat_in,
#                     cols = c("flag_inflamm_factors", "flag_other"), colour=NA,pie_scale = 3) +
#     coord_fixed() +
#     scale_fill_manual(values=color_mat)
# 
# p = p + annotate("text", x = node_cmp_table$xs, y = node_cmp_table$ys,
#                  label = node_cmp_table$cluster_labels, parse = FALSE, size=3, color='grey50',
#                  fontface = 'bold',alpha=0.3)
# 
# #Save out the network as pdf file
# pdf(file = "Figure 3-Inflamm_HPV_PPI_network.pdf",width=8.5*scale, height=11*scale,useDingbats = FALSE) #for inflamm
# print(p)
# dev.off()

```

#Visualize Propagated HPV and SIFs/HPV Preys subnetwork clusters with Viral and SIF Protein overlayed

```{r Add AP-MS links and viral baits to the network ---Uncomment if running for HIV/HPV}
# Initialize nodes and edges tables
edges = fread("~/Codes_for_manuscript_outputs/HPV_Inflamm_edges.txt") # for Inflamm-HPV
#edges = fread("~/Codes_for_manuscript_outputs/HPV_HIV_edges.txt") #for HIV-HPV #for HIV only
edges = data.table(from = edges[,1], to = edges[,2], type = "PC")
nodes_in_edges_all = unique(c(edges$from,edges$to))
nodes_df_clusters = data.frame(nodes = nodes_in_edges_all, cluster = ct[match(nodes_in_edges_all,cc_in$labels,nomatch=F)])

#Extract human proteins that interact only with viral preys--add HIV and HPV or HPV only for the inflamm network
#APMS_links_1 = Krogan_PPI %>% filter(pathogen == "HPV")
APMS_links_1 = Krogan_PPI %>% filter(pathogen == "HIV")

#To include HIV and HPV viral proteins uncomment the following
#APMS_links_1 = Krogan_PPI %>% filter(pathogen == "HPV", pathogen == "HIV")

#Delete HPV-31_
APMS_links_1 = APMS_links_1 %>%
  mutate_at("Bait", str_replace, "HPV-31_", "")

#select column of interest
APMS_links = APMS_links_1[,c("PreyGeneName", "Bait")]
colnames(APMS_links) = c("From", "To")
 
# #convert into a graph object
s1 = rbind(S, APMS_links)

# Make final graph
s = unique(s1)
s = s[as.logical(!rowSums(s=="")),]

# Make and refine graph (only need to do this once and only if needed for your network)
G1 = graph_from_data_frame(s, directed = FALSE, vertices = NULL)
G1 = decompose.graph(G1, max.comps = 2)
G1 = G1[[1]]

# #Load the viral baits columns
viral_prots = as.data.frame(APMS_links_1$Bait)
colnames(viral_prots) = c("viralbaits_added")

#Add the SIFs to the list
SIFs = as.data.frame(inflam_secreted$gene)
colnames(SIFs) = c("viralbaits_added")

#merge the viral proteins and SIFs
viralbaits_added = rbind(viral_prots, SIFs)

viralbaits_added = viralbaits_added %>% distinct(viralbaits_added, .keep_all = TRUE)

#add top hits connections to the nodes table
n=0
G_edges = as.data.table(as_edgelist(G1))
for (i in unique(nodes_df_clusters$cluster)){
  nodes_cluster = as.character(nodes_df_clusters$nodes[nodes_df_clusters$cluster==i])

  # Prepare subnetwork
  G_cluster = G_edges[as.logical(match(G_edges$V1,nodes_cluster,nomatch=F)) & as.logical(match(G_edges$V2,nodes_cluster,nomatch=F)),]
  G_cluster$type = "PC"
  G_cluster$cluster = i

  # Add overall_tophits_addeds
  G_cluster_viralbaits_added = G_edges[
    (as.logical(match(G_edges$V1,viralbaits_added$viralbaits,nomatch=F)) & as.logical(match(G_edges$V2,nodes_cluster,nomatch=F))) | (as.logical(match(G_edges$V1,nodes_cluster,nomatch=F)) & as.logical(match(G_edges$V2,viralbaits_added$viralbaits,nomatch=F)))
    ,]
  G_cluster_viralbaits_added$V1[as.logical(match(G_cluster_viralbaits_added$V1,viralbaits_added$viralbaits,nomatch=F))] = paste(G_cluster_viralbaits_added$V1[as.logical(match(G_cluster_viralbaits_added$V1,viralbaits_added$viralbaits,nomatch=F))],i,sep="_")
  G_cluster_viralbaits_added$V2[as.logical(match(G_cluster_viralbaits_added$V2,viralbaits_added$viralbaits,nomatch=F))] = paste(G_cluster_viralbaits_added$V2[as.logical(match(G_cluster_viralbaits_added$V2,viralbaits_added$viralbaits,nomatch=F))],i,sep="_")

  if (nrow(G_cluster_viralbaits_added)>0){
  G_cluster_viralbaits_added$type = "APMS_added"
  G_cluster_viralbaits_added$cluster = i

  if (n==0){
    G_Final_Subnetworks = rbind(G_cluster,G_cluster_viralbaits_added)
    n=1
  } else {
    G_Final_Subnetworks = rbind(G_Final_Subnetworks,G_cluster,G_cluster_viralbaits_added)
  }

  } else {

    if (n==0){
    G_Final_Subnetworks = G_cluster
    n=1
    } else {
    G_Final_Subnetworks = rbind(G_Final_Subnetworks,G_cluster)
    }

  }
}

# Remove any overall_tophits_addeds that are in network to begin with because they are now replaced by the other overall_tophits_addeds.
G_Final_Subnetworks = G_Final_Subnetworks[!as.logical(match(G_Final_Subnetworks$V1,viralbaits_added$viralbaits,nomatch=F)) & !as.logical(match(G_Final_Subnetworks$V2,viralbaits_added$viralbaits,nomatch=F)), ]

print(length(unique(c(G_Final_Subnetworks$V1,G_Final_Subnetworks$V2))))
print(length(unique(nodes_df_clusters$nodes)))

#save out edge list
fwrite(G_Final_Subnetworks,"Y_final_PPI_network_edges_with_added_viral_baits.txt",sep='\t') # network edgelist

#make node files
nodes = data.frame(nodes = unique(c(as.character(G_Final_Subnetworks$V1),as.character(G_Final_Subnetworks$V2))))
nodes$nodes_raw = gsub("_.*","",nodes$nodes)
nodes$flag_viralbaits_added = as.logical(match(nodes$nodes_raw,viralbaits_added$viralbaits,nomatch=F))

# Add clusters to final nodes file
nodes$cluster = nodes_df_clusters$cluster[match(nodes$nodes_raw,nodes_df_clusters$nodes)]
nodes$cluster_all = nodes$cluster
nodes$cluster_all[grepl("_",nodes$nodes)] = as.numeric(gsub(".*_","",nodes$nodes[grepl("_",nodes$nodes)]))

#Add flags to the nodes
#select host proteins that interact with the viral proteins
HIV_interactors1 = as.data.frame(HIV_interactors)
HPV_interactors1 = as.data.frame(HPV_interactors)
 
# #flag HIV
nodes$flag_HIV_prey = as.logical(match(nodes$nodes_raw,HIV_interactors1$HIV_interactors,nomatch=F))
#
# #Flag HPV
nodes$flag_HPV_prey = as.logical(match(nodes$nodes_raw,HPV_interactors1$gene,nomatch=F))

fwrite(nodes,"Y_final_PPI_network_nodes_and_added_viral_baits.txt",sep='\t') # network nodes
#----------------------------------------------------------------------------------------------------
scale = 1.5
# Load data
G_Final_Subnetworks_1 = fread("~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/Y_final_PPI_network_edges_with_added_viral_baits.txt")

# Load nodes
nodes = fread("~/Codes_for_manuscript_outputs/Network_propagation_genomics_integration_codes/Y_final_PPI_network_nodes_and_added_viral_baits.txt")

# All together
G_edges = G_Final_Subnetworks_1
g = graph_from_data_frame(G_edges)
layout = create_layout(g, layout = 'fr',circular=FALSE)
layout$node_clusters = nodes$cluster_all[match(nodes$nodes,layout$name,nomatch=F)]
node_types = rep("Human protein",nrow(layout))
node_types[as.logical(match(gsub("_.*","",layout$name),viralbaits_added$viralbaits,nomatch=F))] = "viralbaits_added"
layout$node_types = node_types
Y_final[, 2][Y_final[, 2] == 0] <- 0.00001 #replace the pvals = 0 with smaller values
layout$propagation_sig = -log10(Y_final$pvals[match(layout$name,Y_final$nodes)])
layout$propagation_sig[is.na(layout$propagation_sig)]=2
layout[,9] <- rescale(layout[,9],to=c(0.8,1.7))

# Make text annotation table
cmp = components(g)
layout$node_components = cmp$membership
cmps = unique(layout$node_components)
xs = rep(NA,length(cmps))
ys = rep(NA,length(cmps))
cluster_name = rep(NA,length(cmps))
for (i in 1:length(cmps)){
  inds = as.logical(match(layout$node_components,cmps[i],nomatch=F))
  xs[i] = mean(layout$x[inds])
  ys[i] = max(layout$y[inds])
  cluster_name[i]=unique(layout$node_clusters[inds])
}

#ys = ys
node_cmp_table = data.frame(components = cmps, xs = xs, ys = ys,
                            cluster_name = cluster_name, cluster_labels = paste("C",cluster_name,sep=""))

# #--------------------------------------------------------------------------------------------------------------------
# Define node size
node_sizes_pies = layout$node_types
node_sizes_pies[node_sizes_pies=="Human protein"] = 1.5/13
node_sizes_pies[node_sizes_pies=="viralbaits_added"] = 2/13
layout$node_sizes_pies = as.numeric(node_sizes_pies)
layout[is.na(layout)] = 0.2 #replace NA in the node_size_pies

# Create pie mat
data_mat = nodes[,c("flag_HIV_prey","flag_HPV_prey")]
data_mat$flag_HIV_prey = as.numeric(data_mat$flag_HIV_prey)
data_mat$flag_HPV_prey = as.numeric(data_mat$flag_HPV_prey)
data_mat[is.na(data_mat)]=0
data_mat_in = cbind(layout,data_mat)
data_mat_in$node_sizes_pies = layout$propagation_sig * layout$node_sizes_pies
data_mat_in$flag_other = rep(0,nrow(data_mat_in))
data_mat_in$flag_other[rowSums(data_mat_in[,c("flag_HIV_prey","flag_HPV_prey")])==0]=0.25
color_mat = c("red", "green", "darkgrey")

#size_mod = 0.5
p = ggraph(layout) +
    geom_edge_link(aes(color=type, width=1)) +
    scale_edge_width(range = c(0.3, 0.3))+
    scale_edge_color_manual(values = c("PC"="grey", "APMS_added"="red")) +
    scale_color_manual(values = c("red", "darkgrey","red")) +
    scale_shape_manual(values = c(18, 20, 18)) +

    geom_text_repel(aes(x=x,y=y,label=name), point.padding = 0.1, box.padding = 0.15,
                    max.overlaps = Inf,size=1.5,segment.linetype = 'dotted',segment.size=0.2,color="black") +
    theme_classic() +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),
          axis.ticks=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank()) +
    geom_scatterpie(aes(x=x, y=y, r=node_sizes_pies),data = data_mat_in,
                    cols = c("flag_HIV_prey","flag_HPV_prey", "flag_other"),colour="NA",pie_scale = 0) +
    coord_fixed() +
    scale_fill_manual(values=color_mat)

p = p + annotate("text", x = node_cmp_table$xs, y = node_cmp_table$ys,
                 label = node_cmp_table$cluster_labels, parse = FALSE, size=3, color='grey50',
                 fontface = 'bold',alpha=0.3)

#save out network
# pdf(file = "Figure S1-HPV_Inflamm_network_withbaits.pdf",width=8.5*scale, height=11*scale,useDingbats = FALSE)
# print(p)
# dev.off()

# pdf(file = "Figure 5-HPV_HIV_network_withbaits.pdf",width=8.5*scale, height=11*scale,useDingbats = FALSE)
# print(p)
# dev.off()

pdf(file = "Figure 2H-HIV_only_network_withbaits.pdf",width=8.5*scale, height=11*scale,useDingbats = FALSE)
print(p)
dev.off()

```

#Overlaying Top mutated SIFs or HIV preys from CC genomics data from cBioPortal onto the Propagated subnetwork clusters

```{r Prepare network edgelist with top hits from cBioPortal for Cancer Genomics}
# Initialize nodes and edges tables
edges = fread("~/Codes_for_manuscript_outputs/HPV_Inflamm_edges.txt") #Uncomment for inflamm
#edges = fread("~/Codes_for_manuscript_outputs/HPV_HIV_edges.txt") #Uncomment for HIV
edges = data.table(from = edges[,1], to = edges[,2], type = "PC")
nodes_in_edges_all = unique(c(edges$from,edges$to))
nodes_df_clusters = data.frame(nodes = nodes_in_edges_all, cluster = ct[match(nodes_in_edges_all,cc_in$labels,nomatch=F)])

#Load top HIV hits from cBioPortal for Cancer Genomics - only run one dataset at a time
#Top_cbio_hits = fread("~/Codes_for_manuscript_outputs/InPut_datasets/Inflamm_CC_final_hits.txt", header = T) #for SIF hits
Top_cbio_hits = fread("~/Codes_for_manuscript_outputs/InPut_datasets/HIV_CC_final_hits.txt", header = T) #for HIV hits

#Extract and rename columns for grade and stage hits
grade_hits = as.data.frame(Top_cbio_hits$grade_hits)
colnames(grade_hits) = c("gene")

stage_hits = as.data.frame(Top_cbio_hits$stage_hits)
colnames(stage_hits) = c("gene")

#Collapse top hits into a single column add hit connections to the network clusters
overall_tophits_added_1 = rbind(grade_hits, stage_hits)
colnames(overall_tophits_added_1) = c("overall_tophits_added")
overall_tophits_added = subset(overall_tophits_added_1, !duplicated(overall_tophits_added_1[,1])) # delete any duplicates

#add top hits connections to the nodes table
n=0
G_edges = as.data.table(as_edgelist(G))
for (i in unique(nodes_df_clusters$cluster)){
  nodes_cluster = as.character(nodes_df_clusters$nodes[nodes_df_clusters$cluster==i])
  
  # Prepare subnetwork
  G_cluster = G_edges[as.logical(match(G_edges$V1,nodes_cluster,nomatch=F)) & as.logical(match(G_edges$V2,nodes_cluster,nomatch=F)),]
  G_cluster$type = "PC"
  G_cluster$cluster = i
  
  # Add overall_tophits_addeds
  G_cluster_overall_tophits_added = G_edges[
    (as.logical(match(G_edges$V1,overall_tophits_added$overall_tophits_added,nomatch=F)) & as.logical(match(G_edges$V2,nodes_cluster,nomatch=F))) | (as.logical(match(G_edges$V1,nodes_cluster,nomatch=F)) & as.logical(match(G_edges$V2,overall_tophits_added$overall_tophits_added,nomatch=F)))
    ,]
  G_cluster_overall_tophits_added$V1[as.logical(match(G_cluster_overall_tophits_added$V1,overall_tophits_added$overall_tophits_added,nomatch=F))] = paste(G_cluster_overall_tophits_added$V1[as.logical(match(G_cluster_overall_tophits_added$V1,overall_tophits_added$overall_tophits_added,nomatch=F))],i,sep="_") 
  G_cluster_overall_tophits_added$V2[as.logical(match(G_cluster_overall_tophits_added$V2,overall_tophits_added$overall_tophits_added,nomatch=F))] = paste(G_cluster_overall_tophits_added$V2[as.logical(match(G_cluster_overall_tophits_added$V2,overall_tophits_added$overall_tophits_added,nomatch=F))],i,sep="_") 

  if (nrow(G_cluster_overall_tophits_added)>0){
  G_cluster_overall_tophits_added$type = "PC_overall_tophits_added"
  G_cluster_overall_tophits_added$cluster = i
  
  if (n==0){
    G_Final_Subnetworks = rbind(G_cluster,G_cluster_overall_tophits_added)
    n=1
  } else {
    G_Final_Subnetworks = rbind(G_Final_Subnetworks,G_cluster,G_cluster_overall_tophits_added)
  }
  
  } else {
    
    if (n==0){
    G_Final_Subnetworks = G_cluster
    n=1
    } else {
    G_Final_Subnetworks = rbind(G_Final_Subnetworks,G_cluster)
    }
    
  }
}

# Remove any overall_tophits_addeds that are in network to begin with because they are now replaced by the other overall_tophits_addeds.
G_Final_Subnetworks = G_Final_Subnetworks[!as.logical(match(G_Final_Subnetworks$V1,overall_tophits_added$overall_tophits_added,nomatch=F)) & !as.logical(match(G_Final_Subnetworks$V2,overall_tophits_added$overall_tophits_added,nomatch=F)), ]

print(length(unique(c(G_Final_Subnetworks$V1,G_Final_Subnetworks$V2))))
print(length(unique(nodes_df_clusters$nodes)))

#save out edge list
fwrite(G_Final_Subnetworks,"Y_final_network_edges_with_added_CC_hits.txt",sep='\t') # network edgelist

#make node files
nodes = data.frame(nodes = unique(c(as.character(G_Final_Subnetworks$V1),as.character(G_Final_Subnetworks$V2))))
nodes$nodes_raw = gsub("_.*","",nodes$nodes)
nodes$flag_overall_tophits_added = as.logical(match(nodes$nodes_raw,overall_tophits_added$overall_tophits_added,nomatch=F))

# Add clusters to final nodes file
nodes$cluster = nodes_df_clusters$cluster[match(nodes$nodes_raw,nodes_df_clusters$nodes)]
nodes$cluster_all = nodes$cluster
nodes$cluster_all[grepl("_",nodes$nodes)] = as.numeric(gsub(".*_","",nodes$nodes[grepl("_",nodes$nodes)]))

#Add flags to the nodes
#flag HIV
nodes$flag_HIV_prey = as.logical(match(nodes$nodes_raw,HIV_interactors,nomatch=F))

#Flag HPV
nodes$flag_HPV_prey = as.logical(match(nodes$nodes_raw,HPV_interactors,nomatch=F))

#Flag inflamm. markers
nodes$flag_inflamm_factors = as.logical(match(nodes$nodes_raw,inflamm_prots_secreted,nomatch=F))

# #Flag grade hits
top_grade_hits = as.logical(match(nodes$nodes_raw,grade_hits,nomatch=F))
nodes$flag_early_hits[top_grade_hits]="Early_hits"
 
# #Flag stage hits
top_stage_hits = as.logical(match(nodes$nodes_raw,stage_hits,nomatch=F))
nodes$flag_late_hits[top_stage_hits]="Late_hits"

fwrite(nodes,"Y_final_network_nodes_and_added_CC_hits.txt",sep='\t') # network nodes
#---------------------------------------------

# Load data
G_Final_Subnetworks = fread("~/Codes_for_manuscript_outputs/Y_final_network_edges_with_added_CC_hits.txt")

# Load nodes
nodes = fread("~/Codes_for_manuscript_outputs/Y_final_network_nodes_and_added_CC_hits.txt")

# All together
G_edges = G_Final_Subnetworks
g = graph_from_data_frame(G_edges)
layout = create_layout(g, layout = 'fr',circular=FALSE)
layout$node_clusters = nodes$cluster_all[match(nodes$nodes,layout$name,nomatch=F)]
node_types = rep("Human protein",nrow(layout))
node_types[as.logical(match(gsub("_.*","",layout$name),overall_tophits_added$overall_tophits_added,nomatch=F))] = "overall_tophits_added"
layout$node_types = node_types
Y_final[, 2][Y_final[, 2] == 0] <- 0.00001 #replace the pvals = 0 with smaller values
layout$propagation_sig = -log10(Y_final$pvals[match(layout$name,Y_final$nodes)])
layout$propagation_sig[is.na(layout$propagation_sig)]=2
layout[,9] <- rescale(layout[,9],to=c(0.8,1.7))

# Make text annotation table
cmp = components(g)
layout$node_components = cmp$membership
cmps = unique(layout$node_components)
xs = rep(NA,length(cmps))
ys = rep(NA,length(cmps))
cluster_name = rep(NA,length(cmps))
for (i in 1:length(cmps)){
  inds = as.logical(match(layout$node_components,cmps[i],nomatch=F))
  xs[i] = mean(layout$x[inds])
  ys[i] = max(layout$y[inds])
  cluster_name[i]=unique(layout$node_clusters[inds])
}

#ys = ys
node_cmp_table = data.frame(components = cmps, xs = xs, ys = ys, 
                            cluster_name = cluster_name, cluster_labels = paste("C",cluster_name,sep=""))

#--------------------------------------------------------------------------------------------------------------------
# Define node size
node_sizes_pies = layout$node_types
node_sizes_pies[node_sizes_pies=="Human protein"] = 1.5/13
node_sizes_pies[node_sizes_pies=="overall_tophits_added"] = 2/13
layout$node_sizes_pies = as.numeric(node_sizes_pies)

# Create pie mat
data_mat = nodes[,c("flag_HIV_prey","flag_HPV_prey")]
data_mat[data_mat=="Early_hits"]=TRUE
data_mat[data_mat=="Late_hits"]=TRUE
data_mat[data_mat==TRUE]=0.25
data_mat$flag_HIV_prey = as.numeric(data_mat$flag_HIV_prey)
data_mat$flag_HPV_prey = as.numeric(data_mat$flag_HPV_prey)
data_mat$flag_inflamm_factors = as.numeric(data_mat$flag_inflamm_factors)
data_mat$flag_early_hits = as.numeric(data_mat$flag_early_hits)
data_mat$flag_late_hits = as.numeric(data_mat$flag_late_hits)
data_mat[is.na(data_mat)]=0
data_mat_in = cbind(layout,data_mat)
data_mat_in$node_sizes_pies = layout$propagation_sig * layout$node_sizes_pies
data_mat_in$flag_other = rep(0,nrow(data_mat_in))
data_mat_in$flag_other[rowSums(data_mat_in[,c("flag_HIV_prey","flag_HPV_prey", "flag_inflamm_factors", "flag_early_hits", "flag_late_hits")])==0]=0.25

color_mat = c("red","purple", "darkgrey")

#size_mod = 0.5
p = ggraph(layout) + 
    geom_edge_link(aes(color=type, width=1)) +
    scale_edge_width(range = c(0.3, 0.3))+
    scale_edge_color_manual(values = c("PC"="grey", "PC_overall_tophits_added"="blue")) +
    scale_color_manual(values = c("red", "darkgrey","red")) +
    scale_shape_manual(values = c(18, 20, 18)) +

    geom_text_repel(aes(x=x,y=y,label=name), point.padding = 0.1, box.padding = 0.15,
                    max.overlaps = Inf,size=1.5,segment.linetype = 'dotted',segment.size=0.2,color="black") +
    theme_classic() +  
    theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),
          axis.ticks=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank()) +
    geom_scatterpie(aes(x=x, y=y, r=node_sizes_pies),data = data_mat_in,
                    cols = c("flag_HIV_prey","flag_HPV_prey", "flag_inflamm_factors", "flag_early_hits", "flag_late_hits", "flag_other"),colour="NA",pie_scale = 4) +
    coord_fixed() +
    scale_fill_manual(values=color_mat)

p = p + annotate("text", x = node_cmp_table$xs, y = node_cmp_table$ys, 
                 label = node_cmp_table$cluster_labels, parse = FALSE, size=3, color='grey50',
                 fontface = 'bold',alpha=0.3)

#save out network
#pdf(file = "Figure S3A Inflamm_HPV_PPI_network_with_tophits.pdf",width=8.5*scale, height=11*scale,useDingbats = FALSE) #for inflamm
pdf(file = "Figure S3B HIV_HPV_PPI_network_with_tophits.pdf",width=8.5*scale, height=11*scale,useDingbats = FALSE) #for HIV
print(p)
dev.off()

```

#Perform Enrichment Analysis for subnetwork clusters with Connections to the Top Mutated SIFs or HIV preys

```{r Do enrichments per CC_tophit}
# Load data
nodes = fread("~/Codes_for_manuscript_outputs/Y_final_network_nodes_and_added_CC_hits.txt", sep = '\t')

#prepare dataframe for enrichment analysis
for (i in 1:nrow(overall_tophits_added)){
  overall_tophits_added_curr = overall_tophits_added$overall_tophits_added[i]
  cluster_curr = unique(nodes$cluster_all[as.logical(match(nodes$nodes_raw, overall_tophits_added_curr,nomatch=F))])
  genes_curr = nodes$nodes[as.logical(match(nodes$cluster_all,cluster_curr,nomatch=F)) & !nodes$flag_overall_tophits_added]
  if (length(genes_curr)>0){
  if (i==1){
    DF = data.frame(genes=genes_curr, group=overall_tophits_added_curr)
  } else {
    DF = rbind(DF,data.frame(genes=genes_curr, group=overall_tophits_added_curr))
  }
  }
}
# Run enrichment
M=DF
colnames(M)=c("genes","group")
pdf(file = "Figure 3C_Inflamm_HPV_PPI_Pathways_withhits.pdf", width = 8.04, height = 4.78, useDingbats = FALSE) #for Inflamm)
#pdf(file = "Figure S5C_HIV_HPV_PPI_Pathways_withhits.pdf", width = 6.34, height = 3.91, useDingbats = FALSE) #for HIV
enrich_table = do_enrichmap(M,cutHeight=0.8,topN=2,database="Pathways",flag_cluster_col=TRUE,flag_p=FALSE,path_to_databasefiles='~/Codes_for_manuscript_outputs/InPut_datasets/')
dev.off()

#fwrite(enrich_table, "Table S4-Inflamm_overlaid_on_HPV-SIF_network.xls",sep='\t') #for Inflamm
fwrite(enrich_table, "Table S5-HIV_overlaid_on_HIV_HPV_network.xls",sep='\t') #for HIV

```

#Selection of Propagated HPV Prey/SIF/HPV Prey Subnetwork Clusters Relevant to Cervical Cancer

```{r Compare alteration frequencies of network clusters among CC patients}
#Load node tables for HPV/SIFs propagation --- Uncomment to run for HPV/SIFs
HPV_inflamm_subnetworks1 = fread("~/Codes_for_manuscript_outputs/InPut_datasets/HPV_inflamm_HIV_nodes.txt", header = T)
HPV_inflamm_subnetworks = HPV_inflamm_subnetworks1 %>% dplyr::select(nodes, clusters)
setnames(HPV_inflamm_subnetworks, "nodes", "gene_symbol")
# 
# #NB: EXTRACT THE ALTERATION FREQUENCIES (Mutation, CNA, mRNA and protein expression alterations) OF ALL NODES FOR CERVICAL CANCERS STUDIES (Pancancer & Firehose Legacy)
# 
# #PANCANCER and FIREHOSE MERGED
CC_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_inflamm_clusters_pancancer.tsv"))
CC_TCGA1$freq1 = CC_TCGA1$num_samples_altered / 297 * 100

CC_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_inflamm_clusters_firehouse.tsv"))
CC_TCGA2$freq2 = CC_TCGA2$num_samples_altered / 310 * 100

CC_combined = CC_TCGA1 %>%
     left_join(CC_TCGA2, by='gene_symbol')
CC_combined$Alteration_frequency_in_TCGA = (CC_combined$freq1 + CC_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
CC_genet_clusters = CC_combined %>%
  left_join(HPV_inflamm_subnetworks, by='gene_symbol')

#compare cervical cancer mut frequencies across clusters
CC_genet_clusters = na.omit(CC_genet_clusters)
p = ggplot(CC_genet_clusters, aes(x = fct_reorder(as_factor(clusters), Alteration_frequency_in_TCGA, .desc = FALSE), y = Alteration_frequency_in_TCGA))+
  geom_boxplot(outlier.shape = NA, width = 0.7)+
  scale_y_continuous(name="Alteration frequency in TCGA (%)", limits = c(0, 30)) +
  xlab("Clusters") +
  theme_classic(base_size = 12)+
  theme(legend.position="none")+
  coord_flip()

pdf(file = "Figure 2B-CC_genomics_HPV_inflamm_19clusters.pdf",width=2.96, height=3.52,useDingbats = FALSE)
print(p)
dev.off()

#---------------------------------------------------------------------------------------------------
#Load node tables for HPV/HIV propagation --- uncomment to run for HIV/HPV preys 
HPV_HIV_subnetworks1 = fread("~/Codes_for_manuscript_outputs/InPut_datasets/HPV_HIV_nodes.txt", header = T)
HPV_HIV_subnetworks = HPV_HIV_subnetworks1 %>% dplyr::select(nodes, clusters)
setnames(HPV_HIV_subnetworks, "nodes", "gene_symbol")

#PANCANCER and FIREHOSE MERGED
CC_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_HIV_clusters_pancancer.tsv"))
CC_TCGA1$freq1 = CC_TCGA1$num_samples_altered / 297 * 100

CC_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_HIV_clusters_firehouse.tsv"))
CC_TCGA2$freq2 = CC_TCGA2$num_samples_altered / 310 * 100
# 
CC_combined = CC_TCGA1 %>%
     left_join(CC_TCGA2, by='gene_symbol')
CC_combined$Alteration_frequency_in_TCGA = (CC_combined$freq1 + CC_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
CC_genet_clusters = CC_combined %>%
  left_join(HPV_HIV_subnetworks, by='gene_symbol')

#compare cervical cancer mut frequencies across clusters
CC_genet_clusters = na.omit(CC_genet_clusters)
# 
p = ggplot(CC_genet_clusters, aes(x = fct_reorder(as_factor(clusters), Alteration_frequency_in_TCGA, .desc = FALSE), y = Alteration_frequency_in_TCGA))+
  geom_boxplot(outlier.shape = NA, width = 0.7)+
  scale_y_continuous(name="Alteration frequency in TCGA (%)") +
  xlab("Clusters") +
  theme_classic() +
  theme(legend.position="none")+
  coord_flip()
pdf(file = "Figure S2B-CC_genomics_HPV_HIV33clusters.pdf",width=4.47, height=4.95,useDingbats = FALSE)
print(p)
dev.off()

#Select top 5 altered clusters for subsequent analyses

```

#Comparison of the alteration frequencies of the 5 selected clusters from previous chunks across different cancers

```{r SIF/HPV networsk: Compare genetic alteration for selected clusters across different cancers---Uncomment this chunk if running SIF/HPV}
#NB: EXTRACT THE ALTERATION FREQUENCIES (Mutation, CNA, mRNA and protein expression alterations) OF ALL NODES FOR COMMON CANCERS (lung, liver, cervical cancer, breast, colorectal and stomach) AMONG WOMEN FROM CBIOPORTAL DATABASE

#Load alteration frequencies for the SIF/HPV network nodes
#PANCANCER and FIREHOUSE MERGED
#cervical cancer
CC_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_inflamm_clusters_pancancer.tsv"))
CC_TCGA1$freq1 = CC_TCGA1$num_samples_altered / 297 * 100

CC_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_inflamm_clusters_firehouse.tsv"))
CC_TCGA2$freq2 = CC_TCGA2$num_samples_altered / 310 * 100

CC_combined = CC_TCGA1 %>%
     left_join(CC_TCGA2, by='gene_symbol')
CC_combined$Alteration_frequency_in_TCGA = (CC_combined$freq1 + CC_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
CC_genet_clusters = CC_combined %>%
  left_join(HPV_inflamm_subnetworks, by='gene_symbol')
CC_genet_clusters$Cancer_type = rep("Cervical", nrow(CC_genet_clusters))

#Lung
LG_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LUC_all_inflamm_clusters_pancancer.tsv"))
LG_TCGA1$freq1 = LG_TCGA1$num_samples_altered / 566 * 100

LG_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LUC_all_inflamm_clusters_firehouse.tsv"))
LG_TCGA2$freq2 = LG_TCGA2$num_samples_altered / 586 * 100

LG_combined = LG_TCGA1 %>%
     left_join(LG_TCGA2, by='gene_symbol')
LG_combined$Alteration_frequency_in_TCGA = (LG_combined$freq1 + LG_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
LG_genet_clusters = LG_combined %>%
  left_join(HPV_inflamm_subnetworks, by='gene_symbol')
LG_genet_clusters$Cancer_type = rep("Lung", nrow(LG_genet_clusters))

#Breast
BR_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/BR_all_inflamm_clusters_pancancer.tsv"))
BR_TCGA1$freq1 = BR_TCGA1$num_samples_altered / 1084 * 100

BR_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/BR_all_inflamm_clusters_firehouse.tsv"))
BR_TCGA2$freq2 = BR_TCGA2$num_samples_altered / 1108 * 100

BR_combined = BR_TCGA1 %>%
     left_join(BR_TCGA2, by='gene_symbol')
BR_combined$Alteration_frequency_in_TCGA = (BR_combined$freq1 + BR_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
BR_genet_clusters = BR_combined %>%
  left_join(HPV_inflamm_subnetworks, by='gene_symbol')
BR_genet_clusters$Cancer_type = rep("Breast", nrow(BR_genet_clusters))

#Colorectum
CR_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/COL_all_inflamm_clusters_pancancer.tsv"))
CR_TCGA1$freq1 = CR_TCGA1$num_samples_altered / 594 * 100

CR_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/COL_all_inflamm_clusters_firehouse.tsv"))
CR_TCGA2$freq2 = CR_TCGA2$num_samples_altered / 640 * 100

CR_combined = CR_TCGA1 %>%
     left_join(CR_TCGA2, by='gene_symbol')
CR_combined$Alteration_frequency_in_TCGA = (CR_combined$freq1 + CR_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
CR_genet_clusters = CR_combined %>%
  left_join(HPV_inflamm_subnetworks, by='gene_symbol')
CR_genet_clusters$Cancer_type = rep("Colorectal", nrow(CR_genet_clusters))

#Liver
LV_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LVC_all_inflamm_clusters_pancancer.tsv"))
LV_TCGA1$freq1 = CC_TCGA1$num_samples_altered / 372 * 100

LV_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LVC_all_inflamm_clusters_firehouse.tsv"))
LV_TCGA2$freq2 = CC_TCGA2$num_samples_altered / 379 * 100
LV_combined = LV_TCGA1 %>%
     left_join(LV_TCGA2, by='gene_symbol')
LV_combined$Alteration_frequency_in_TCGA = (LV_combined$freq1 + LV_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
LV_genet_clusters = LV_combined %>%
  left_join(HPV_inflamm_subnetworks, by='gene_symbol')
LV_genet_clusters$Cancer_type = rep("Liver", nrow(LV_genet_clusters))

#Stomach
STOM_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/STOM_all_inflamm_clusters_pancancer.tsv"))
STOM_TCGA1$freq1 = STOM_TCGA1$num_samples_altered / 440 * 100

STOM_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/STOM_all_inflamm_clusters_firehouse.tsv"))
STOM_TCGA2$freq2 = STOM_TCGA2$num_samples_altered / 478 * 100

STOM_combined = STOM_TCGA1 %>%
     left_join(STOM_TCGA2, by='gene_symbol')
STOM_combined$Alteration_frequency_in_TCGA = (STOM_combined$freq1 + STOM_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
STOM_genet_clusters = STOM_combined %>%
  left_join(HPV_inflamm_subnetworks, by='gene_symbol')
STOM_genet_clusters$Cancer_type = rep("Stomach", nrow(STOM_genet_clusters))

#-----------------------------------------------------------------------------------------------------------

#Merged df with all the clusters, genes and muts
D_super_genet = rbind(CC_genet_clusters,LG_genet_clusters,LV_genet_clusters,BR_genet_clusters,CR_genet_clusters, STOM_genet_clusters)

#Extract rows corresponding to Rho GTPpase pathway clusters
Rhogtpase_subnetwork_selected = dplyr::filter(D_super_genet, clusters %in% c("15"))
Rhogtpase_subnetwork_selected$Pathways = rep("Rho gtpase-15", nrow(Rhogtpase_subnetwork_selected))

#Extract rows corresponding to DNA damage response pathway clusters
DNAdam_subnetwork_selected = dplyr::filter(D_super_genet, clusters %in% c("13"))
DNAdam_subnetwork_selected$Pathways = rep("DNA damage response-13", nrow(DNAdam_subnetwork_selected))

#Extract rows corresponding to PI3K pathway clusters
PI3K_subnetwork_selected = dplyr::filter(D_super_genet, clusters %in% c("9", "10", "14"))
PI3K_subnetwork_selected$Pathways = rep("PI3K-AKT-9,10,14", nrow(PI3K_subnetwork_selected))

#merge the 3 pathway data frames
top5pathways = rbind(Rhogtpase_subnetwork_selected, DNAdam_subnetwork_selected, PI3K_subnetwork_selected)

top5pathway_median_super = aggregate(top5pathways$Alteration_frequency_in_TCGA,
by = list(top5pathways$Cancer_type, top5pathways$Pathways),
FUN = function(x){median(x,na.rm=T)})

#plot
p = ggplot(top5pathway_median_super, aes(x = Group.2, y = x, label=Group.1)) +
  geom_boxplot(color = "black") +
  geom_point(size=3) +
  scale_y_continuous(name="Alteration frequency in TCGA (%)") +
  xlab("Pathways") +
   geom_text_repel(check_overlap = TRUE,
            position=position_jitter(width=0.2))+
  theme_classic()

pdf(file = "Figure 2E-topclusters_selected.pdf",width=3.74, height=3.86,useDingbats = FALSE)
print(p)
dev.off()

#Plot bars comparing genet.alt for PIK3CA across the 6 cancers
#combine datasets for ploting
D_super = rbind(CC_genet_clusters,LG_genet_clusters,LV_genet_clusters,STOM_genet_clusters,BR_genet_clusters,CR_genet_clusters)
D_super_sub = subset(D_super, gene_symbol == "PIK3CA")

#Prepare dataframe for plotting barplots with error bars
pancancer_PIK3CA_dat = D_super_sub %>% dplyr::select(freq1, Cancer_type)
colnames(pancancer_PIK3CA_dat) = c("freq", "Cancer_type")

firehose_PIK3CA_dat = D_super_sub %>% dplyr::select(freq2, Cancer_type)
colnames(firehose_PIK3CA_dat) = c("freq", "Cancer_type")

PIK3CA_alter_cancers_new <- PIK3CA_alter_cancers %>%
  group_by(Cancer_type) %>%
  mutate(mean_freq = mean(freq)) %>%
  mutate(se = std.error(freq))

p3 = ggplot(PIK3CA_alter_cancers_new, aes(x = fct_reorder(as_factor(Cancer_type), mean_freq, .desc=TRUE), y=mean_freq, fill = Cancer_type)) + 
  geom_bar(stat="identity", width=0.4,color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean_freq, ymax=mean_freq+se), width=.2,
                 position=position_dodge(.75))+
  scale_fill_brewer(palette = "Paired")+
  ylab("PIK3CA alteration frequency in TCGA (%)")+
  xlab("Cancer type")+
  theme_bw()+
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
      )+
    theme(strip.background = element_rect(fill="white"))+
  guides(x =  guide_axis(angle = 90))+
  theme(legend.position="none")

pdf(file = "Figure 2F-PIK3CA_genomics_multiple_cancers.pdf",width=2.83, height=3.81,useDingbats = FALSE)
print(p3)
dev.off()

#Load top 5 mutated genes in CC from cBioportal
top5muts_cc = fread("~/Codes_for_manuscript_outputs/InPut_datasets/top_5_mut_genes_CC.txt")

top5muts_cc_new <- top5muts_cc %>%
  group_by(gene) %>%
  mutate(mean_freq = mean(mut_freq)) %>%
  mutate(se = std.error(mut_freq))

p4 = ggplot(top5muts_cc_new, aes(x = fct_reorder(as_factor(gene), mean_freq, .desc=TRUE), y=mean_freq, fill = gene)) + 
  geom_bar(stat="identity", width=0.4,color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean_freq, ymax=mean_freq+se), width=.2,
                 position=position_dodge(.75))+
  scale_fill_brewer(palette = "Paired")+
  ylab("Mut. freq in CC (%)")+
  xlab("")+
  theme_bw()+
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
      )+
    theme(strip.background = element_rect(fill="white"))+
  guides(x =  guide_axis(angle = 90))+
  theme(legend.position="none")
 
pdf(file = "Figure 2F_inset-top5_mut_genes_CC.pdf",width=1.7, height=1.5,useDingbats = FALSE)
print(p4)
dev.off()


#####################################################################################################
#Ugandan PIK3CA mutations in HIV+ vs HIV- tumors
#Load dataset
UG_data = fread("~/Codes_for_manuscript_outputs/InPut_datasets/Ugandan_PIK3CA_muts.txt")

p = ggplot(UG_data, aes(x = status, y = frequency, fill=status))+
  geom_bar(stat='identity', width = 0.6)+
  scale_y_continuous(name="Number of tumors (%)") +
  xlab("") +
  theme_classic()+
  scale_fill_brewer(palette="Paired") +
  theme(legend.position="none")+
  #theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))#+
  coord_flip()

pdf(file = "Figure 3D-PIK3CA_muts_vs_HIV_status.pdf",width=2.74, height=1.04,useDingbats = FALSE)
print(p)
dev.off()

```

```{r HIV/HPV networrk: Compare genetic alteration for selected clusters across different cancers --- Uncomment this chunk if running HPV/HIV}
#NB: EXTRACT THE ALTERATION FREQUENCIES (Mutation, CNA, mRNA and protein expression alterations) OF ALL NODES FOR COMMON CANCERS (lung, liver, cervical cancer, breast, colorectal and stomach) AMONG WOMEN FROM CBIOPORTAL DATABASE

#Load alteration frequencies for the HIV/HPV network nodes
#PANCANCER and FIREHOUSE MERGED
#cervical cancer
CC_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_HIV_clusters_pancancer.tsv"))
CC_TCGA1$freq1 = CC_TCGA1$num_samples_altered / 297 * 100

CC_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/CC_all_HIV_clusters_firehouse.tsv"))
CC_TCGA2$freq2 = CC_TCGA2$num_samples_altered / 310 * 100

CC_combined = CC_TCGA1 %>%
     left_join(CC_TCGA2, by='gene_symbol')
CC_combined$Alteration_frequency_in_TCGA = (CC_combined$freq1 + CC_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
CC_genet_clusters = CC_combined %>%
  left_join(HPV_HIV_subnetworks, by='gene_symbol')
CC_genet_clusters$Cancer_type = rep("Cervical", nrow(CC_genet_clusters))
#
# #Lung
LG_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LUC_all_HIV_clusters_pancancer.tsv"))
LG_TCGA1$freq1 = LG_TCGA1$num_samples_altered / 566 * 100

LG_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LUC_all_HIV_clusters_firehouse.tsv"))
LG_TCGA2$freq2 = LG_TCGA2$num_samples_altered / 586 * 100

LG_combined = LG_TCGA1 %>%
     left_join(LG_TCGA2, by='gene_symbol')
LG_combined$Alteration_frequency_in_TCGA = (LG_combined$freq1 + LG_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
LG_genet_clusters = LG_combined %>%
  left_join(HPV_HIV_subnetworks, by='gene_symbol')
LG_genet_clusters$Cancer_type = rep("Lung", nrow(LG_genet_clusters))
#
# #Breast
BR_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/BR_all_HIV_clusters_pancancer.tsv"))
BR_TCGA1$freq1 = BR_TCGA1$num_samples_altered / 1084 * 100

BR_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/BR_all_HIV_clusters_firehouse.tsv"))
BR_TCGA2$freq2 = BR_TCGA2$num_samples_altered / 1108 * 100

BR_combined = BR_TCGA1 %>%
     left_join(BR_TCGA2, by='gene_symbol')
BR_combined$Alteration_frequency_in_TCGA = (BR_combined$freq1 + BR_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
BR_genet_clusters = BR_combined %>%
  left_join(HPV_HIV_subnetworks, by='gene_symbol')
BR_genet_clusters$Cancer_type = rep("Breast", nrow(BR_genet_clusters))
#
# #Colorectum
CR_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/COL_all_HIV_clusters_pancancer.tsv"))
CR_TCGA1$freq1 = CR_TCGA1$num_samples_altered / 594 * 100

CR_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/COL_all_HIV_clusters_firehouse.tsv"))
CR_TCGA2$freq2 = CR_TCGA2$num_samples_altered / 640 * 100

CR_combined = CR_TCGA1 %>%
     left_join(CR_TCGA2, by='gene_symbol')
CR_combined$Alteration_frequency_in_TCGA = (CR_combined$freq1 + CR_combined$freq2) / 2

# #Extract genetic alter frequency for subnetwork nodes
CR_genet_clusters = CR_combined %>%
  left_join(HPV_HIV_subnetworks, by='gene_symbol')
CR_genet_clusters$Cancer_type = rep("Colorectal", nrow(CR_genet_clusters))
#
# # #Liver
LV_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LVC_all_HIV_clusters_pancancer.tsv"))
LV_TCGA1$freq1 = CC_TCGA1$num_samples_altered / 372 * 100

LV_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/LVC_all_HIV_clusters_firehouse.tsv"))
LV_TCGA2$freq2 = CC_TCGA2$num_samples_altered / 379 * 100
LV_combined = LV_TCGA1 %>%
     left_join(LV_TCGA2, by='gene_symbol')
LV_combined$Alteration_frequency_in_TCGA = (LV_combined$freq1 + LV_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
LV_genet_clusters = LV_combined %>%
  left_join(HPV_HIV_subnetworks, by='gene_symbol')
LV_genet_clusters$Cancer_type = rep("Liver", nrow(LV_genet_clusters))
#
# #Stomach
STOM_TCGA1 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/STOM_all_HIV_clusters_pancancer.tsv"))
STOM_TCGA1$freq1 = STOM_TCGA1$num_samples_altered / 440 * 100

STOM_TCGA2 = clean_names(fread("~/Codes_for_manuscript_outputs/InPut_datasets/STOM_all_HIV_clusters_firehouse.tsv"))
STOM_TCGA2$freq2 = STOM_TCGA2$num_samples_altered / 478 * 100

STOM_combined = STOM_TCGA1 %>%
     left_join(STOM_TCGA2, by='gene_symbol')
STOM_combined$Alteration_frequency_in_TCGA = (STOM_combined$freq1 + STOM_combined$freq2) / 2

#Extract genetic alter frequency for subnetwork nodes
STOM_genet_clusters = STOM_combined %>%
  left_join(HPV_HIV_subnetworks, by='gene_symbol')
STOM_genet_clusters$Cancer_type = rep("Stomach", nrow(STOM_genet_clusters))
#
# #-----------------------------------------------------------------------------------------------------------
#
# #Merged df with all the clusters, genes and muts
D_super_HIVgenet = rbind(CC_genet_clusters,LG_genet_clusters,LV_genet_clusters,BR_genet_clusters,CR_genet_clusters, STOM_genet_clusters)

#Extract rows for selected networks that have Rhogtpase pathways
splicing_subnetwork_selected = dplyr::filter(D_super_HIVgenet, clusters %in% c("16", "23", "30", "32"))
splicing_subnetwork_selected$Pathways = rep("mRNA splicing-cluster 16, 23, 30, 32", nrow(splicing_subnetwork_selected))

#Extract rows for selected networks that have DNA damage response pathways
Hdr_ssa_subnetwork_selected = dplyr::filter(D_super_HIVgenet, clusters %in% c("33"))
Hdr_ssa_subnetwork_selected$Pathways = rep("HDR-ssa-33", nrow(Hdr_ssa_subnetwork_selected))
#
# #merge the 3 pathway dfs
top5HIVpathways = rbind(splicing_subnetwork_selected, Hdr_ssa_subnetwork_selected)

top5HIVpathway_median_super = aggregate(top5HIVpathways$Alteration_frequency_in_TCGA,
by = list(top5HIVpathways$Cancer_type, top5HIVpathways$Pathways),
FUN = function(x){median(x,na.rm=T)})
#
# #plot
p = ggplot(top5HIVpathway_median_super, aes(x = Group.2, y = x, label=Group.1)) +
  geom_boxplot(color = "black") +
  geom_point(col="forestgreen", size=3) +
  scale_y_continuous(name="Alteration frequency in TCGA (%)") +
  xlab("Pathways") +
   geom_text_repel(check_overlap = TRUE,
            position=position_jitter(width=0.2))+
    coord_flip()+
  theme(legend.position="none")+
  theme_classic()

pdf(file = "Figure S2E-topclusters_selected.pdf",width=7.9, height=2.94,useDingbats = FALSE)
print(p)
dev.off()

```

```{r Network for PI3K, SIFs and HPV proteins }
#Krogan PPI
Krogan_PPI = read.table("~/Codes_for_manuscript_outputs/InPut_datasets/Krogan_HIV_HPV_AP-MS_interactome.txt", header = T, sep = '\t')

# Load HPV-prey network (data.frame with From, To columns)
HPV_APMS = Krogan_PPI %>% filter(pathogen == "HPV")

#Delete HPV-31_
#HPV_APMS = HPV_APMS %>%
  #mutate_at("Bait", str_replace, "HPV-31_", "")

#select column of interest
HPV_prey_int = HPV_APMS[,c("PreyGeneName", "Bait")]
colnames(HPV_prey_int) = c("From", "To")

# Load your base network (Pathway Commons with From, To columns)
KD_pathwaycommons = fread('D:/Charles_PhD_Stuff/R_scripts_notebooks/Special_files/PathwayCommons_Clean.txt') #Download from PathwayCommons

#select column of interest
PC = KD_pathwaycommons[,c("From", "To")]

# Merge your HPV-prey network into your base network (rbind)
HPV_prey_PC = rbind(HPV_prey_int, PC)

# Save out your network to tsv file
fwrite(HPV_prey_PC, "STRING.gt600.allRec.tsv", sep = "\t", col.names=FALSE)

#Prepare base network of PC and HPV PPIs
BN = fread("STRING.gt600.allRec.tsv", header = FALSE)
# Make final graph
S = unique(BN)
S = S[as.logical(!rowSums(S=="")),]

# Make and refine graph (only need to do this once and only if needed for your network)
G = graph_from_data_frame(S, directed = FALSE, vertices = NULL)
G = decompose.graph(G, max.comps = 2)
G = G[[1]]

#List of top50 signif PI3K genes + SIFs + HPV genes
Y_final = fread("~/Codes_for_manuscript_outputs/InPut_datasets/output_propagations_diffusion_individual.txt")
Y.wide = reshape2::dcast(Y_final,nodes~dataset,value.var="pvals")
Y.wide$sig[rowSums(Y.wide[,c("HPV_preys","Secreted_inflamm")]) < 0.05]="YES" #convergence of HPV preys vs SIFs
Y.wide$sigsum = rowSums(Y.wide[c("HPV_preys","Secreted_inflamm")]) #calculate rowsum
Y.wide_clean = subset(Y.wide, sig=="YES")

#Merge sig propagated genes and PI3K genes
PIK3genes = as.data.frame(PIK3genes_clean$value)
colnames(PIK3genes) = ("nodes")

selected_PI3K = PIK3genes %>%
  left_join(Y.wide_clean, by="nodes")

prop_df_sorted2 = arrange(selected_PI3K, sigsum)
selected_sig_PI3K = as.data.frame(prop_df_sorted2[1:100, "nodes"]) #select top 100 most significant genes
colnames(selected_sig_PI3K) = c("nodes")

SIFs = as.data.frame(inflamm_prots_secreted)
colnames(SIFs) = c("nodes")

HPV_prots = HPV_prots_clean
colnames(HPV_prots) = c("nodes")

#Network list
all_genes = rbind(selected_sig_PI3K, SIFs, HPV_prots)

# Make subgraph
SN = induced_subgraph(G, all_genes$nodes)

# Select only fully connected components of subnetwork
SN = igraph::simplify(SN)
components = igraph::clusters(SN, mode="weak")
biggest_cluster_id = which.max(components$csize)
vert_ids = V(SN)[components$membership == biggest_cluster_id]
SN = igraph::induced_subgraph(SN, vert_ids)

edge_list = as.data.frame(as_edgelist(SN, names = TRUE))

#Select the nodes that interact with SIFs
edge_list$flag_SIFs1 = as.logical(match(edge_list$V1,inflamm_prots_secreted,nomatch=F))
edge_list$flag_SIFs2 = as.logical(match(edge_list$V2,inflamm_prots_secreted,nomatch=F))

#Clean the rows without SIFs
edge_list$KK = 0
#add UP
edge_list$KK[edge_list$flag_SIFs1 == TRUE | edge_list$flag_SIFs2 == TRUE] = "1"
edge_list_clean = subset(edge_list, KK==1)
edge_list_clean = edge_list_clean[,c("V1","V2","KK")]

#Save out network edges file with cluster annotation
fwrite(as.data.frame(edge_list_clean),file="PI3K-SIF_edges_all.xls",sep='\t') #visualize in cytoscape

```

```{r END}

```
